---
title: "Sentiment Analysis"
author: "Lei Fan"
date: "7/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Sentiment Analysis in R

This follows Data Flair's [tutorial](https://data-flair.training/blogs/data-science-r-sentiment-analysis-project/), with minor modifications.

```{r load_libraries, include=FALSE}
library(tidytext)
library(janeaustenr)
library(stringr)
library(tidytext)
library(dplyr)
library(tidyr)
library(ggplot2)
library(reshape2)
library(wordcloud)
```

## Sentiments of Words in Pride & Prejudice

How many words in the book *Pride & Prejudice* have positive/negative sentiment, according to Bing's sentiment analysis?

First, let's take a look at word count...

```{r pp_word_count, message=FALSE}

pp_words <- austen_books() %>% 
              filter(book == 'Pride & Prejudice') %>% 
              mutate(linenumber=row_number(), 
                     chapter=cumsum(str_detect(text, 
                                               regex('^chapter [\\dxlciv]', 
                                                     ignore_case=TRUE)))) %>% 
              unnest_tokens(word, text)

pp_word_count <- pp_words %>% count(word, sort=TRUE)

head(pp_word_count)

```

What about word count when filtered by positive and negative sentiments?

```{r pp_sentiment, message=FALSE}

sentiments <- get_sentiments('bing')

pp_word_sentiments <- pp_word_count %>% inner_join(sentiments)

head(pp_word_sentiments)

```


What is the sentiment per chapter?

```{r pp_sentiment_per_chapter, message=FALSE}

pp_chapter_sentiments <- pp_words %>% 
                          inner_join(sentiments) %>% 
                          count(index = chapter, sentiment) %>% 
                          pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
                          mutate(sentiment = positive - negative)

pp_chapter_sentiments %>% ggplot(aes(index, sentiment)) + geom_bar(stat = 'identity', color = 'blue')

```

What are the most common words in the book, and what are their sentiments?

```{r common_word_sentiments, message=FALSE}

pp_word_sentiments %>% 
  filter(n > quantile(n, 0.97)) %>% 
  mutate(n = ifelse(sentiment == 'negative', -n, n)) %>% 
  mutate(word=reorder(word, n)) %>% 
  ggplot(aes(word, n, fill=sentiment)) + 
    geom_col() + 
    coord_flip() + 
    ylab('Sentiment Score')

```

Let's see the same info in a word cloud.

```{r wordcloud, message=FALSE}

pp_word_sentiments %>% 
  acast(word ~ sentiment, value.var='n', fill = 0) %>% 
  comparison.cloud(colors=c('red', 'blue'), 
                   max.words=100)

```