---
title: "Statistical Inference Course Project"
author: "Lei Fan"
date: "12/28/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This assignment demonstrates the Central Limit Theorem with the exponential distribution by showing that the means of many samples from an exponential distribution are normally distributed.

## Simulations

We will demonstrate this with 1000 random samplings of 40 samples each from the expontential distribution with a rate parameter (lambda) of 0.2.

```{r simulations}
exps = NULL
n = 40
runs = 1000
lambda = 0.2
for (i in 1 : runs) exps = c(exps, rexp(n, lambda))
exps = matrix(exps, runs, n)
mns = apply(exps, 1, mean)
vars = apply(exps, 1, var)
```

## Plots

We can see the original samples from the exponential distribution looks very exponential:

```{r exponential, echo=FALSE, fig.height=3, fig.width=4, fig.align="center"}
hist(exps, main="Sample distribution")
```

But the mean looks normally distributed:

```{r mean, echo=FALSE, fig.height=3, fig.width=4, fig.align="center"}
hist(mns, main="Mean of the sample distribution")
```

The mean of the sample means is:

```{r mean_of_sample_means}
mean(mns)
```

The variance of the sample means is:

```{r variance_of_sample_means}
var(mns)
```

As expected, the mean of the sample mean is very close to the theoretical mean, `1/lambda = 5`, and the variance is very close to the theoretical variance divided by n, `(1/lambda)^2/n = 0.625` since, according to the CLT, the sample means is approximately normal with a mean equal to the population mean and a variance equal to the population variance divided by the sample count.

We can calculate the 95% confidence interval of the mean:
```{r confidence_interval}
t.test(mns)$conf.int
```

And see that the theoretical mean does indeed lie within the 95% confidence interval calculated from the sample means.