---
title: "Tooth Growth Data Analysis"
author: "Lei Fan"
date: "12/28/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
options(width=80)
```

## Overview

This project looks at the Tooth Growth dataset, part of the R dataset package, and determines what effect the dose and delivery method of vitamin C has on tooth growth in guinea pigs, if any.

## Experiment Setup

60 guinea pigs were evenly divided into 6 groups of 10 guinea pigs each: 3 groups were given vitamin C as orange juice in doses of 0.5, 1, and 2 mg/day, respectively, and the remaining 3 groups were given vitamin C as ascorbic acid in doses of 0.5, 1, and 2 mg/day, respectively.

## Data Analysis

```{r boxplot, echo=FALSE}
boxplot(len ~ supp + dose, ToothGrowth, xlab = "Delivery method and dosage", ylab = "Tooth Length", main="Tooth Growth in Guinea Pigs from Vitamin C")
```

It appears that higher doses of vitamin C is positively correlated with greater tooth growth, and that, at doses below 2 mg/day, orange juice promotes greater tooth growth than ascorbic acid.

Let's look at the means and standard variations of tooth growth of each group:

```{r summary}
tbl_df(ToothGrowth) %>% group_by(supp, dose) %>% summarize(mean = mean(len), sd = sd(len))
```

Is it true that, on average, higher doses of vitamin C is positively correlated with greater tooth growth?

```{r linear_model}
summary(lm(len ~ dose, ToothGrowth))
```

We see a very small p-value (well below an alpha of 0.05), and a relatively strong R-squared value, indicating that there does seem to be a positive correlation between the two factors.

Is orange juice more effective than ascorbic acid?

At all doses (alpha = 0.05):

```{r comp_all_doses}
t.test(subset(ToothGrowth, supp=='OJ')$len, subset(ToothGrowth, supp=='VC')$len, alternative = "g", var.equal = FALSE)
```

We get a p-value of 0.03, thus we reject the null hypothesis in favor of the hypothesis that orange juice is generally more effective than ascorbic acid at promoting tooth growth.

This is even more visible at lower doses:

```{r comp_lower_doses}
t.test(subset(ToothGrowth, supp=='OJ' & dose==0.5)$len, subset(ToothGrowth, supp=='VC' & dose==0.5)$len, alternative = "g", var.equal = FALSE)
t.test(subset(ToothGrowth, supp=='OJ' & dose==1.0)$len, subset(ToothGrowth, supp=='VC' & dose==1.0)$len, alternative = "g", var.equal = FALSE)
```

The p-values are very small, thus we reject the null hypothesis in favor of the alternative hypothesis that orange juice is more effective than ascorbic acid at promoting tooth growth.